### 2015-01-31 Projects/DelCo_2014/Report/Hose_Tests/DelCo_Hose_Tests_2014.tex
DelCo_2014: Created outline for hose test report and edited hose test report
DelCo_2014: Added content to the hose test report and changed title of outline

\headerB{
Delco 2014
}

@@

\chapter{Introduction}
\label{chap:Introduction}
NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire. The studies have resulted in guidance to the fire service regarding ventilation tactics. However ventilation tactics alone will not result in the complete extinguishment of the fire, fire suppression with hose streams are needed.

@@

\subsection{Hose Stream Patterns}
\label{sec:Hose_Stream_Patterns}

\subsection{Application Patterns}
\label{Application_Patterns}

@@

\subsection{Construction}
\label{sec:Construction}
Two structures, shown in Fig.~\ref{fig:struct_pics}, were constructed for the experiments. Each structure contained a ground level floor composed of concrete walls. The West Structure contained a second level.

@@

\section{Instrumentation}
\label{sec:Instrumentation}
A schematic plan overview of the instrumentation in the West Structure referenced in this report is shown in Fig.~\ref{fig:west_instrumentation}. There is a discussion of uncertainties for each measurement below in Section~\ref{sec:Uncertainty}. Differential pressure transducers connected to bi-directional velocity probes (BDPs) were used to measure gas velocity at specific points throughout the structure ~\cite{McCaffrey:Combustion_and_Flame}. Each set of BDPs contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95 m below the soffit of the corresponding doorway. Fig.~\ref{fig:BDPs} shows two sets of BDPs, A5 and A6, which are located at each double door on the first floor of the West Structure.

@@

\section{Uncertainty}
\label{sec:Uncertainty}
There are different components of uncertainty in the length, differential pressure, and gas velocity reported in this report. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means [46]. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Multiplying the combined standard uncertainty by a coverage factor of two results in the expanded uncertainty which correspond to a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from instrument specifications. For other components, such as differential pressure, past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions,
instrumentation array locations, and fan placement were made with a hand held laser measurement
device which has an accuracy of $\pm$6.0 mm over a range of 0.61 m to 15.3 m [47]. However, conditions affecting the measurement, such as levelness of the device, yield an
estimated uncertainty of $\pm$0.5\% for measurements in the 2.0 m to 10.0 m range. Steel measuring tapes with a resolution of $\pm$0.5 mm were used to locate individual sensors within a measurement array and to measure and position the furniture. Some issues, such as "soft" edges on the upholstered furniture, result in an estimated total expanded uncertainty of $\pm$1.0\%.

Differential pressure reading uncertainty components were derived from pressure transducer instrument specifications and previous experience with pressure transducers. The transducers were factory calibrated and the zero and span of each was checked in the laboratory prior to the experiments yielding an accuracy of $\pm$1\% [52]. The total expanded uncertainty was estimated at 10\%. Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouple are co-located with the probe. The estimated total expanded uncertainty for velocity in these experiments is $\pm$18\%.

%46. Taylor, B.N., and Kuyatt, C.E., “Guidelines for Evaluating and Expressing the Uncertainty of NIST Measurement Results”, National Institute of Standards and Technology, Gaithersburg. MD., NIST TN 1297, January 1993. 
%47. Stanley Hand Tools, User Manual TLM 100, 1000 Stanley Drive, New Britain, CT 06053. 
%52. Setra Model 264 Very Low Pressure Transducer Data Sheet Rev E. Setra Systems, Boxborough, MA., December 2002.

@@

\subsection{Water Flow Experimental Procedure}
\label{sec:Water_Flow_Procedure}

Two sets of experiments involving application patterns, Tests 18 and 19, were conducted. The two sets varied in that the first floor south side door was open for Test 18 (Fig~\ref{fig:test_18_plan}) and closed for Test 19 (Fig.~\ref{fig:test_19_plan}).

>>>>

\headerB{
Impact of Hose Streams on Air Flows inside a Structure
}

@@

\chapter{Introduction}
\label{chap:Introduction}
NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not result in the complete extinguishment of the fire; fire suppression with hose streams are needed.

@@

@@

\subsection{Construction}
\label{sec:Construction}
Two structures, shown in Fig.~\ref{fig:struct_pics}, were constructed for the experiments. Each structure contained a ground level floor composed of walls surrounded by concrete blocks. The West Structure contained a second level.

@@

\section{Instrumentation}
\label{sec:Instrumentation}
Schematic plan overviews of the instrumentation in the East and West Structures are shown in Fig.~\ref{fig:east_instrumentation} and Fig.~\ref{fig:west_instrumentation}, respectively. There is a discussion of uncertainties for each measurement below in Section~\ref{sec:Uncertainty}. Gas velocity was measured at various doorways using differential pressure transducers connected to bi-directional velocity probes~\cite{McCaffrey:Combustion_and_Flame}. Each set of BDPs contained 8 probes located at distances of 0.08 m, 0.34 m, 0.61 m, 0.88 m, 1.15 m, 1.42 m, 1.68 m, and 1.95 m below the soffit of the corresponding doorway. Fig.~\ref{fig:BDPs} shows a picture of two sets of BDPs, A5 and A6, located at each double door on the first floor of the West Structure. A single thermocouple was attached to each bi-directional probe. The thermocouples used with the bi-directional probes are exposed-bead, Chromel-Alumel (type K) with a 1.0 mm (0.04 in) diameter. Starting with the exposed bead, the thermocouple wire was sheathed in a 3.2 mm (0.13 in) diameter Inconel shield, 0.76 m (2.5 ft) in length.

%\begin{figure}[!ht]
%\includegraphics[trim=0cm 0cm 0.75cm 4.5cm, clip=true, width=6in]{../../Drawings/PDFs/With_Instrumentation/East_Test_Structure_Devices_Hose_Test}
%\caption[Location of Instrumentation in East Structure]{Location of Instrumentation in East Structure}
%\label{fig:east_instrumentation}
%\end{figure}

@@

\section{Uncertainty}
\label{sec:Uncertainty}
There are different components of uncertainty in the length, differential pressure, and gas velocity reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means ~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions,
instrumentation array locations, and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of $\pm6.0$ mm (0.24 in) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft)~\cite{Stanley}. However, conditions affecting the measurement, such as levelness of the device, yield an
estimated uncertainty of $\pm$0.5\% for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range. Steel measuring tapes with a resolution of $\pm$0.5 mm were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm1.6$ mm (0.06 in) for 9.1 m (30 ft) tapes and $\pm6.4$ mm (0.25 in) for 30.5 m (100 ft) tapes \cite{NIST_Manual_44}. Some issues, such as "soft" edges on the upholstered furniture, result in an estimated total expanded uncertainty of $\pm$1.0\%.

Bi-directional probes and single thermocouples were used to measure the velocity. The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above. Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design \cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments estimated to be $\pm18$\%.

Water flowrate was measured with a pressure and flow meter combination shown in Fig. \ref{fig:flow_meter}. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0–4.1 MPa (0-600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\%  for the pressure sensor \cite{Akron}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at $\pm10$\%.

%\begin{figure}[!ht]
%\includegraphics[width=6in]{../../Figures/flow_meter}
%\caption[Picture of Flowrate Meter]{Pressure and flow meter combination used to measure water flowrate}
%\label{fig:flow_meter}
%\end{figure}

In the following sections, the measurements will be presented in graphic and tabular form. In the graphs, an error bar will represent the estimated uncertainty of the measurement. In the tables, the uncertainty will be included in the table of as part of the caption.

@@

\subsection{Water Flow Experimental Procedure}
\label{sec:Water_Flow_Procedure}
A total of four sets of experiments were conducted to observe the impact of using different hose streams and application patterns on the ventilation in the West Structure.

Two sets of experiments involving application patterns, Tests 18 and 19, were conducted. The two sets varied in that the first floor south side door was open for Test 18 (Fig.~\ref{fig:test_18_plan}) and closed for Test 19 (Fig.~\ref{fig:test_19_plan}).
